---
title: "Assignment 1"
author: "Giridhar Manoharan"
date: "2/2/2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---



```{r setup, include=FALSE}
library(igraph)
library(ggplot2)
require(scales)
library(knitr)
```

## Question 1
```{r question1, results='asis',error=F,warning=F}
internetGraph = read_graph("./as-22july06/as-22july06.gml", format = "gml")
politicalGraph = read_graph("./polblogs/polblogs.gml", format = "gml")
neuralNetGraph = read_graph("./celegansneural/celegansneural.gml", format = "gml")
erdosRenyi01 = erdos.renyi.game(2000, 0.01)
erdosRenyi005 = erdos.renyi.game(2000, 0.005)
erdosRenyi0025 = erdos.renyi.game(2000, 0.0025)
networks = data.frame(Network = character(6), Type = character(6), n = integer(6), m = integer(6), c_strong = integer(6), c_weak = integer(6), d = integer(6), l = double(6), L = integer(6), cc_local = double(6), cc_global = double(6), stringsAsFactors = FALSE)

#political blogs
networks$Network[1] = "Political blogs"
networks$Type[1] = if(is_directed(politicalGraph) == TRUE) "DIRECTED" else "UNDIRECTED"
networks$n[1] = vcount(politicalGraph)
networks$m[1] = ecount(politicalGraph)
networks$c_strong[1] = count_components(politicalGraph, mode = "strong")
networks$c_weak[1] = count_components(politicalGraph, mode = "weak")
networks$d[1] = max(degree(politicalGraph))
networks$l[1] = mean_distance(politicalGraph)
networks$L[1] = diameter(politicalGraph)
networks$cc_local[1] = transitivity(politicalGraph, type = "localaverageundirected")
networks$cc_global[1] = transitivity(politicalGraph, type = "globalundirected")

#neural network
networks$Network[2] = "Neural network"
networks$Type[2] = if(is_directed(neuralNetGraph) == TRUE) "DIRECTED" else "UNDIRECTED"
networks$n[2] = vcount(neuralNetGraph)
networks$m[2] = ecount(neuralNetGraph)
networks$c_strong[2] = count_components(neuralNetGraph, mode = "strong")
networks$c_weak[2] = count_components(neuralNetGraph, mode = "weak")
networks$d[2] = max(degree(neuralNetGraph))
networks$l[2] = mean_distance(neuralNetGraph)
networks$L[2] = diameter(neuralNetGraph)
networks$cc_local[2] = transitivity(neuralNetGraph, type = "localaverageundirected")
networks$cc_global[2] = transitivity(neuralNetGraph, type = "globalundirected")

#internet
networks$Network[3] = "Internet"
networks$Type[3] = if(is_directed(internetGraph) == TRUE) "DIRECTED" else "UNDIRECTED"
networks$n[3] = vcount(internetGraph)
networks$m[3] = ecount(internetGraph)
networks$c_strong[3] = count_components(internetGraph)
networks$c_weak[3] = NA
networks$d[3] = max(degree(internetGraph))
networks$l[3] = mean_distance(internetGraph)
networks$L[3] = diameter(internetGraph)
networks$cc_local[3] = transitivity(internetGraph, type = "localaverageundirected")
networks$cc_global[3] = transitivity(internetGraph, type = "globalundirected")

#Erdos Renyi 0.01
networks$Network[4] = "Erdos Renyi 0.01"
networks$Type[4] = if(is_directed(erdosRenyi01) == TRUE) "DIRECTED" else "UNDIRECTED"
networks$n[4] = vcount(erdosRenyi01)
networks$m[4] = ecount(erdosRenyi01)
networks$c_strong[4] = count_components(erdosRenyi01)
networks$c_weak[4] = NA
networks$d[4] = max(degree(erdosRenyi01))
networks$l[4] = mean_distance(erdosRenyi01)
networks$L[4] = diameter(erdosRenyi01)
networks$cc_local[4] = transitivity(erdosRenyi01, type = "localaverageundirected")
networks$cc_global[4] = transitivity(erdosRenyi01, type = "globalundirected")

#Erdos Renyi 0.005
networks$Network[5] = "Erdos Renyi 0.005"
networks$Type[5] = if(is_directed(erdosRenyi005) == TRUE) "DIRECTED" else "UNDIRECTED"
networks$n[5] = vcount(erdosRenyi005)
networks$m[5] = ecount(erdosRenyi005)
networks$c_strong[5] = count_components(erdosRenyi005)
networks$c_weak[5] = NA
networks$d[5] = max(degree(erdosRenyi005))
networks$l[5] = mean_distance(erdosRenyi005)
networks$L[5] = diameter(erdosRenyi005)
networks$cc_local[5] = transitivity(erdosRenyi005, type = "localaverageundirected")
networks$cc_global[5] = transitivity(erdosRenyi005, type = "globalundirected")

#Erdos Renyi 0.0025
networks$Network[6] = "Erdos Renyi 0.0025"
networks$Type[6] = if(is_directed(erdosRenyi0025) == TRUE) "DIRECTED" else "UNDIRECTED"
networks$n[6] = vcount(erdosRenyi0025)
networks$m[6] = ecount(erdosRenyi0025)
networks$c_strong[6] = count_components(erdosRenyi0025)
networks$c_weak[6] = NA
networks$d[6] = max(degree(erdosRenyi0025))
networks$l[6] = mean_distance(erdosRenyi0025)
networks$L[6] = diameter(erdosRenyi0025)
networks$cc_local[6] = transitivity(erdosRenyi0025, type = "localaverageundirected")
networks$cc_global[6] = transitivity(erdosRenyi0025, type = "globalundirected")

colnames(networks) = c("Network", "Type", "Nodes", "Links", "Connected components (Strongly connected - directed)", "Weakly connected components", "Maximum degree", "Average path length", "Diameter", "Average local clustering coefficient", "Global clustering coefficient")

kable(networks, format = "markdown")
```

## Question 2
```{r question2}
#political blogs
plot(degree.distribution(politicalGraph), pch=19, cex=1.2, col="orange", xlab="node degree (k)", ylab = "probability P(k)", main = "Degree distribution of Political blogs network")
lines(degree_distribution(politicalGraph))

#neural network
plot(degree.distribution(neuralNetGraph), pch=19, cex=1.2, col="orange", xlab="node degree (k)", ylab = "probability P(k)", main = "Degree distribution of Neural network")
lines(degree_distribution(neuralNetGraph))

#internet
plot(degree.distribution(internetGraph), pch=19, cex=1.2, col="orange", xlab="node degree (k)", ylab = "probability P(k)", main = "Degree distribution of Internet")
lines(degree_distribution(internetGraph))

#Erdos Renyi 0.01
plot(degree.distribution(erdosRenyi01), pch=19, cex=1.2, col="orange", xlab="node degree (k)", ylab = "probability P(k)", main = "Degree distribution of Erdos-Renyi G(2000,0.01)")
lines(degree_distribution(erdosRenyi01))

#Erdos Renyi 0.005
plot(degree.distribution(erdosRenyi005), pch=19, cex=1.2, col="orange", xlab="node degree (k)", ylab = "probability P(k)", main = "Degree distribution of Erdos-Renyi G(2000,0.005)")
lines(degree_distribution(erdosRenyi005))

#Erdos Renyi 0.0025
plot(degree.distribution(erdosRenyi0025), pch=19, cex=1.2, col="orange", xlab="node degree (k)", ylab = "probability P(k)", main = "Degree distribution of Erdos-Renyi G(2000,0.0025)")
lines(degree_distribution(erdosRenyi0025))
```

**Observations**
From the above degree distribution graphs we can observe that real world networks like "Political blogs", "Neural networks", and "Internet" have power law distribution. Whereas, random graphs generated using Erdos-Renyi model have poisson (bell-shaped) distribution.

## Question 3
```{r question3}
#political blogs
d = path.length.hist(politicalGraph)$res
plDist = data.frame(pathLength = integer(length(d)), freq = integer(length(d)))
plDist$pathLength = 1:length(d)
plDist$freq = d
ggplot(plDist, aes(x=pathLength, y=freq)) + geom_bar(stat = "identity", fill = "steelblue") + labs(title="Path length distribution - Political blogs network", x="path length", y="number of paths") + scale_y_continuous(labels = comma)

#neural network
d = path.length.hist(neuralNetGraph)$res
plDist = data.frame(pathLength = integer(length(d)), freq = integer(length(d)))
plDist$pathLength = 1:length(d)
plDist$freq = d
ggplot(plDist, aes(x=pathLength, y=freq)) + geom_bar(stat = "identity", fill = "steelblue") + labs(title="Path length distribution - Neural network", x="path length", y="number of paths") + scale_y_continuous(labels = comma)

#internet
d = path.length.hist(internetGraph)$res
plDist = data.frame(pathLength = integer(length(d)), freq = integer(length(d)))
plDist$pathLength = 1:length(d)
plDist$freq = d
ggplot(plDist, aes(x=pathLength, y=freq)) + geom_bar(stat = "identity", fill = "steelblue") + labs(title="Path length distribution - Internet", x="path length", y="number of paths") + scale_y_continuous(labels = comma)

#Erdos Renyi 0.01
d = path.length.hist(erdosRenyi01)$res
plDist = data.frame(pathLength = integer(length(d)), freq = integer(length(d)))
plDist$pathLength = 1:length(d)
plDist$freq = d
ggplot(plDist, aes(x=pathLength, y=freq)) + geom_bar(stat = "identity", fill = "steelblue") + labs(title="Path length distribution - Erdos-Renyi G(2000,0.01)", x="path length", y="number of paths") + scale_y_continuous(labels = comma)

#Erdos Renyi 0.005
d = path.length.hist(erdosRenyi005)$res
plDist = data.frame(pathLength = integer(length(d)), freq = integer(length(d)))
plDist$pathLength = 1:length(d)
plDist$freq = d
ggplot(plDist, aes(x=pathLength, y=freq)) + geom_bar(stat = "identity", fill = "steelblue") + labs(title="Path length distribution - Erdos-Renyi G(2000,0.005)", x="path length", y="number of paths") + scale_y_continuous(labels = comma)

#Erdos Renyi 0.0025
d = path.length.hist(erdosRenyi0025)$res
plDist = data.frame(pathLength = integer(length(d)), freq = integer(length(d)))
plDist$pathLength = 1:length(d)
plDist$freq = d
ggplot(plDist, aes(x=pathLength, y=freq)) + geom_bar(stat = "identity", fill = "steelblue") + labs(title="Path length distribution - Erdos-Renyi G(2000,0.0025)", x="path length", y="number of paths") + scale_y_continuous(labels = comma)

```

**Observations**
From the above path length distributions, we can observe that all the six networks have bell-shaped path length distribution. We can also observe that in erdos renyi random graphs, as we increase the probability of an edge between any two nodes, path lengths take lower values. For lower probabilities of an edge between two nodes, path lengths with higher values can be found.


## Question 4
I have chosen power grid real world network here to do network analysis. It is an undirected, unweighted network representing the topology of the Western States Power Grid of the United States. The nodes represent power plants and substations, and links represent transmission (high voltage) lines.
```{r question4}
powerGridGraph = read_graph("./power/power.gml", format = "gml")
networks = data.frame(Network = character(1), Type = character(1), n = integer(1), m = integer(1), c = integer(1), d = integer(1), l = double(1), L = integer(1), cc_local = double(1), cc_global = double(1), stringsAsFactors = FALSE)
networks$Network[1] = "Power grid network"
networks$Type[1] = if(is_directed(powerGridGraph) == TRUE) "DIRECTED" else "UNDIRECTED"
networks$n[1] = vcount(powerGridGraph)
networks$m[1] = ecount(powerGridGraph)
networks$c[1] = count_components(powerGridGraph)
networks$d[1] = max(degree(powerGridGraph))
networks$l[1] = mean_distance(powerGridGraph)
networks$L[1] = diameter(powerGridGraph)
networks$cc_local[1] = transitivity(powerGridGraph, type = "localaverageundirected")
networks$cc_global[1] = transitivity(powerGridGraph, type = "globalundirected")
colnames(networks) = c("Network", "Type", "Nodes", "Links", "Connected components", "Maximum degree", "Average path length", "Diameter", "Average local clustering coefficient", "Global clustering coefficient")
kable(networks, format = "markdown")

plot(degree.distribution(powerGridGraph), pch=19, cex=1.2, col="orange", xlab="node degree (k)", ylab = "probability P(k)", main = "Degree distribution of Power grid network")
lines(degree_distribution(powerGridGraph))

d = path.length.hist(powerGridGraph)$res
plDist = data.frame(pathLength = integer(length(d)), freq = integer(length(d)))
plDist$pathLength = 1:length(d)
plDist$freq = d
ggplot(plDist, aes(x=pathLength, y=freq)) + geom_bar(stat = "identity", fill = "steelblue") + labs(title="Path length distribution - Power grid network", x="path length", y="number of paths") + scale_y_continuous(labels = comma)
```
